Epigraphy and onomastics in the Hesperia databank

### 2016-01-20 source/24/Epigraphy and onomastics in Hesperia databank-template.tex
updated paper 24
including bibliography

\item[c] Authoritative work. In a traditional corpus, editors impose their knowledge in a large range of features, and thus
they interfere in the utility of other views, to the extend of depreciating them too much. When it comes to an
epigraphic corpus that is written in a wide-known language, such as Latin, this auctoritas becomes particularly
powerful regarding the reading of the epigraphic text, its checking being not possible all the time for users,
particularly for those who are not experts on the subject. But when it comes to fragmentary languages, with very low or
even no language deciphering at all, and with many doubts on decoding, the truth is that the influence of the editor is
extremely high and involves some risks that they do not always come to terms with successfully.

@@

Thanks to the search engine, all these points can then be redistributed into smaller groups to draw linguistic areas or
isoglosses, on the basis, for instance, of the attestation of a well-known anthroponymic element or the dispersal area
of significant phonetic features. In the following maps it is displayed, in the first place, the attestation area of
names containing the element biur, which clearly corresponds to the Iberian area, that is, the non indo-European half
of the Iberian peninsula; and, in the second place, it is displayed the dispersal area of the name Tancinus, which
clearly corresponds to the indo-European part.

\begin{figure}[!bp]
\centering
 \includegraphics[width=\columnwidth]{EpigraphyandonomasticsinHesperiadatabanktemplate-img008.jpg}
\caption{Dispersal area of the anthroponymic element biur}
\label{fig:}
\end{figure}

@@

A similar consideration might also apply, for instance, to another important document in the Palaeohispanic epigraphic
landscape, namely the III Botorrita Bronze plaque, written in Celtiberian language and script but containing a
considerable amount of foreign names, such as Latin or Greek.\footnote{ Vid. Unterman’s proposal in MLH IV, K.1.3. }
This could be the case for markos (Marcus), bolora (Flora), kinbiria (Cimbria), antiokos (Antiochus), bilinos
(Philinus), bilonikos (Philonicus), tais (Thais), tiokenes (Diogenes), and so on. We could finally mention, in this
same regard, the Gaullish names observed in some Iberian inscriptions from southern France: aśetile (Adsedilus, CIL III
5373), tesile (Tessillus, CIL III 14368.28), uaśile (Vassil(l)us, CIL XII 2286), katubaŕe (Catumarus, CIL 3, 4263),
etc.

>>>>

\item[c] Authoritative work. In a traditional corpus, editors impose their knowledge in a large range of features, and thus
they interfere in the utility of other views, to the extent of depreciating them too much. When it comes to an
epigraphic corpus that is written in a wide-known language, such as Latin, this auctoritas becomes particularly
powerful regarding the reading of the epigraphic text, its checking being not possible all the time for users,
particularly for those who are not experts on the subject. But when it comes to fragmentary languages, with very low or
even no language deciphering at all, and with many doubts on decoding, the truth is that the influence of the editor is
extremely high and involves some risks that they do not always come to terms with successfully.

@@

Thanks to the search engine, all these points can then be redistributed into smaller groups to draw linguistic areas or
isoglosses, on the basis, for instance, of the attestation of a well-known anthroponymic element or the dispersal area
of significant phonetic features. In the following maps it is displayed, in the first place, the attestation area of
names containing the element \emph{biur}, which clearly corresponds to the Iberian area, that is, the non indo-European half
of the Iberian peninsula; and, in the second place, it is displayed the dispersal area of the name Tancinus, which
clearly corresponds to the indo-European part.

\begin{figure}[!bp]
\centering
 \includegraphics[width=\columnwidth]{EpigraphyandonomasticsinHesperiadatabanktemplate-img008.jpg}
\caption{Dispersal area of the anthroponymic element \emph{biur}}
\label{fig:}
\end{figure}

@@

A similar consideration might also apply, for instance, to another important document in the Palaeohispanic epigraphic
landscape, namely the III Botorrita Bronze plaque, written in Celtiberian language and script but containing a
considerable amount of foreign names, such as Latin or Greek.\footnote{ Vid. Unterman’s proposal in MLH IV, K.1.3. }
This could be the case for \emph{markos} (\emph{Marcus}), \emph{bolora} (\emph{Flora}), \emph{kinbiria} (\emph{Cimbria}), \emph{antiokos} (\emph{Antiochus}), \emph{bilinos}
(\emph{Philinus}), \emph{bilonikos} (\emph{Philonicus}), \emph{tais} (\emph{Thais}), \emph{tiokenes} (\emph{Diogenes}), and so on. We could finally mention, in this
same regard, the Gaullish names observed in some Iberian inscriptions from southern France: \emph{aśetile} (\emph{Adsedilus}, CIL III
5373), \emph{tesile} (\emph{Tessillus}, CIL III 14368.28), \emph{uaśile} (\emph{Vassil(l)us}, CIL XII 2286), \emph{katubaŕe} (\emph{Catumarus}, CIL 3, 4263),
etc.
